---
title: "Trajectory analysis for late dynamics manuscript"
author: "Rita Grunbreg"
date: "11/09/2021"
output: html_document
---
Metadata for bamboo_community and stentor_community.csv

DATE = observation date

JAR = original microcosm coding where B = Bamboo (assemblage A) and S = Stentor (assemblage B) and R = replicate

JAR2 = microcosms coded numerically

TRT= treatment where 1 = +I/+R, 2 = +I/-R, 3 = -I/-R, 4 = -I/+R, 5 = -R for assemblage A, 
and 6 = +I/+R, 7 = +I/-R, 8= -I/-R, 9= -I/+R, and 10 = -R for assemblage B; 
I = invader, R = resident, + = evolved - = naive

Columns following TRT are the species present in each assemblage. 

For the invader in the control treatment (-R), the period indicates a value of NA from SAS because these treatments lack the invader by design

Prorodon niveus (a resident) failed to establish in the +R/-I treatment and was therefore analyzed separately in all previous analyses (in analyses that drop that treatment)

DATA ARE MEAN COUNTS PER ML

E. daidaleos invaded Assemblage A. 

```{r setup,  results='hide', message=FALSE, warning=FALSE}
rm(list = ls())  # clear Rs brain 
library(readr)   # read csv file
library(tidyr)   # tidying data
library(dplyr)   # manipulating df
library(ggplot2) # graphing
library(ggpubr)  # nice package for figures
library(vegan)   # nmds 
library(trend)   # mann-kendell trend test
library(lmerTest) #mixed model to be used in piecewise model 
library(mvabund)
###############################################################################################################
bamboo_community <- read_csv("C:/Users/grunberg/Documents/GitHub/Late-dynamics/data/bamboo_community.csv")
set.seed(123456789)
```
Formatting csv files 

Includes making plotting theme and setting color palette
Adding interpretable treatments into data
Calculating days between protists counts 


*for caras job talk use these colors
 Darkest green: +I/+R (Coevolved), RBG 61,91,80, '#3D5B50'
 "darker" medium green: +I/-R (Evolved invaders), RBG 92, 136, 120 "#5C8878"
 medium green: -I/+R (Evolved residents), RBG 126, 156, 76 "#7E9C4C" 
 light green: -I/-R (Naïve invasion)RBG 181, 202, 146 "#B5CA92"
 
```{r bamboo community formatting}
rita_theme <-  theme_classic(base_size = 15)+
  theme(legend.title = element_text(size=7, face="bold"),
        legend.text = element_text(size=6.5),
        legend.position=c(0.15,0.18),
        axis.line = element_line(colour = 'black', size = 0.75),
        axis.ticks = element_line(colour = "black", size = 0.75),
        axis.text.x = element_text(size=12,  colour = "black"), 
        axis.text.y = element_text(size=12,  colour = "black"), # bold
        axis.title.y = element_text(size=14, colour = "black")
  )

#color palette 
colorpal <- c('#d7191c','#fdae61','#abdda4','#2b83ba') # color for job
colorpal2 <- c('#d7191c','#fdae61','#abdda4','#2b83ba', '#f0f0f0') # color for MS

unique(bamboo_community$JAR)

Treatments <- data.frame(TRT = seq(1:5),
                         treat = c("+I/+R", "+I/-R", "-I/-R", "-I/+R", "-R" ),
                         plot.trt = c("Coevolved", "Evolved invader", "Naive invasion",
                                      "Evolved residents", "Uninvaded"),
                         invader_evol = c("evolved", "evolved", "naive", "naive", "naive"),
                         contrast_invader = c("1", "1", "-1", "-1", "-1"),
                         resident_evol=c("evolved", "naive", "naive", "evolved", "naive"),
                         contrast_resident = c("1", "-1", "-1", "1", "NA"),
                         contrast_interaction =c ("-1", "1","-1", "1", "NA"),
                        invasion= c("invaded", "invaded", "invaded", "invaded", 'uninvaded')
                        )

bamboo_community <- bamboo_community %>%
  mutate(survey_date = as.Date(as.character(DATE), format="%m/%d/%Y"))%>%
  group_by(JAR) %>% arrange(survey_date) %>%
  mutate(BETWEEN0=as.numeric(difftime(survey_date,lag(survey_date,1))),
         BETWEEN=ifelse(is.na(BETWEEN0),0,BETWEEN0),
         FIRST=cumsum(as.numeric(BETWEEN)))%>%
  select(-BETWEEN0)%>%
  drop_na(TRT)


bamboo_community <- merge(bamboo_community, Treatments, by =c("TRT"))

#check <-bamboo_community %>% filter(plot.trt == 'Uninvaded')

#bamboo_community %>% group_by(plot.trt) %>% 
 # mutate(Euplotes_daidaleos1= as.numeric(Euplotes_daidaleos1)) %>%
  #summarise(maxE = max(Euplotes_daidaleos1),
   #         maxEp = min(Euplotes_patella))

#dataframe for resident only analysis
bamboo_community_res <- bamboo_community %>%   
  select(-c(Euplotes_daidaleos1))  # remove invader 

#dataframe for evolutionary history analysis
bamboo_community <- bamboo_community %>%   
  filter(!(treat == "-R")) %>%# remove -R (uninvaded control) not a good comparison bc lacks invader E. daidaleos
  mutate(Euplotes_daidaleos1= as.numeric(Euplotes_daidaleos1)) 

```
Now formatting the stentor commmunity aka assemblage B

Added time of the surveys in days to data frame and treatment names for plotting 

I established patterns of similarity using non-metric multi-dimensional scaling on Bray-Curtis distances. Below is the code to run the ordination and extract relevant information for plotting and analysis 

Note that in Assemblage A we excluded Prorodon from the analysis because it did not establish

```{r NMDS assemblage A}
# create community matrix and envi data matrix 
bamb_mat<-bamboo_community[,c(6:10,12)] # matrix of protist species
bamb_mat_treat<-bamboo_community[,c(1:5, 15:19)] # matrix of treatments

 # global nmds
set.seed(123456789)
 ordO_pt<-metaMDS(bamb_mat, distance="bray", trymax = 500, autotransform=FALSE)
 ordO_pt

 #extract info for plotting
 #add species names to ordination
  species <- c("B. americanum", "S. teres", "E. patella", "E. daidaleos", "Lecane sp.", #"Prorodon",
               "P. bursaria")
  hcoordO<-as.data.frame(scores(ordO_pt, display="sites"))#extracts coordinates for plot
  pcoordO<-scores(ordO_pt, display="species")#extracts coordinates for parasite vectors
  pcoordO <-as.data.frame(pcoordO)
  pcoordO$species <- species
 
  #make dataframe with NMDS coordinates and treatment info 
 bamboo_all <- bind_cols(bamb_mat_treat, hcoordO)
 #write.csv(bamboo_all, "NMDS_coordinates_bamboo_assemblage.csv")
```

Calculate the centroids for analysis and plotting community trjaectories 
```{r bamboo NMDS centroids}
cent.bamboo <- bamboo_all %>%
  group_by(plot.trt, Survey) %>%
  summarise_at(c("NMDS1", 'NMDS2'), mean)
```

Graphic of all the data from the NMDS

```{r graphic_NMDS}
topp1<-max(bamboo_all[,11:12]) #determines maximum and minimum values for the plot axes
bott1<-min(bamboo_all[,11:12]) #I draw from both data sets because I make the graph sqaure and

#jpeg(filename="AssemblageA_alldata.jpeg", width=180, height=180, units="mm", bg="white", res=300)
ggplot(bamboo_all, aes(x= NMDS1, y=NMDS2))+
  geom_hline(yintercept = 0, lty=2, color="grey") +
  geom_vline(xintercept = 0, lty=2, color="grey") +
  geom_point( size =5,  pch=21, aes(fill=plot.trt))+
  xlab("NMDS 1") +
  ylab("NMDS 2") +
  # xlim(c( bott-0.1, topp+0.1)) +
  #ylim(c( bott-0.1, topp+0.1)) +
  geom_text( aes(x=NMDS1,y=NMDS2, label=Survey), fontface =1, colour = "black" , size=4) + # plant species label
  geom_text(data=pcoordO, aes(x=NMDS1,y=NMDS2, label=species), fontface =2, colour = "black" , size=3) + # species label
  scale_fill_manual(values =colorpal)+
  guides(fill=FALSE)+
  theme_pubr() +
  theme(legend.text = element_text(size=8), legend.box = "horizontal",
        legend.title = element_text(size=9, face="bold"),
        legend.position=c(0.13,0.85)) +
  guides(colour = guide_legend(override.aes = list(size=70)))+
  theme(plot.title = element_text(hjust = 0.5, size=15, face="bold")) + # guides(shape=FALSE)+ 
  facet_wrap(~plot.trt,
             ncol = 2, nrow = 3)+
  guides(fill=FALSE, color=FALSE)+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        strip.text = element_text(size = 15)) # add black square panel around graphic
#dev.off()
```
Centroids to visualize the trajectories 

```{r graphic_nmds_centroid assemblage A}
topp<-max(cent.bamboo[,3:4]) #determines maximum and minimum values for the plot axes
bott<-min(cent.bamboo[,3:4]) #I draw from both data sets because I make the graph sqaure and

# #jpeg(filename="AssemblageA_trajectory.jpeg", width=180, height=180, units="mm", bg="white", res=300)
jpeg(filename="AssemblageA_trajectory.jpeg", width=180, height=180, units="mm", bg="white", res=300)
ggplot(data= cent.bamboo, aes(x= NMDS1, y=NMDS2))+
  geom_hline(yintercept = 0, lty=2, color="grey") +
  geom_vline(xintercept = 0, lty=2, color="grey") +
  geom_path(color="#525252", lwd=1)+
  geom_point( size =6, aes(x= NMDS1, y=NMDS2, fill= plot.trt ),
              data=cent.bamboo, pch=21)+
  xlab("NMDS 1") +
  ylab("NMDS 2") +
  xlim(c(-1.5, 1)) +
  ylim(c( -1.0, 1)) +
  scale_shape_manual(values=c(21,22,23,24))+
  scale_fill_manual(values =colorpal)+
  geom_text( aes(x=NMDS1,y=NMDS2, label=Survey), fontface =2, colour = "black" , size=5) + # plant species label
  geom_text(aes(x=NMDS1,y=NMDS2, label=Survey), fontface =1, colour = "black" , size=4.5) + # plant species label
  geom_text(data=pcoordO, aes(x=NMDS1,y=NMDS2, label=species), fontface =1, colour = "black" , size=3) + # species label
  theme_pubr() +
  theme(legend.text = element_text(size=8), legend.box = "horizontal",
        legend.title = element_text(size=9, face="bold"),
        legend.position=c(0.13,0.85)) +
  guides(colour = guide_legend(override.aes = list(size=70)))+
  theme(plot.title = element_text(hjust = 0.5, size=15, face="bold")) + # guides(shape=FALSE)+
  facet_wrap(~plot.trt,
             ncol = 2, nrow = 2)+
  guides(fill=FALSE, color=FALSE)+
  ggtitle("Assemblage A")+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        strip.text = element_text(size = 15))
dev.off()

# #dev.off()

```

Below calculates distance between communities for each treatment over time
this will test for convergence between communities 

```{r trajectory}
###################################################################################################
# distance between treatment centroids: how dissimilar are the trajectories btwn groups 
###################################################################################################
dist.pt <- function(x1, y1, x2, y2) sqrt((x2-x1)^2 + (y2-y1)^2) # finds distance between points 

cent.bamboo <- bamboo_all %>%
  group_by(plot.trt, Survey) %>%
  summarise_at(c("NMDS1", 'NMDS2'), mean) 

trt_list <- unique(cent.bamboo$plot.trt)
dist_btw_trt <- data.frame()

for (i in trt_list){
  for(j in trt_list){
    
    treat <- i
    treat2 <- j 
    
    trt_one <- cent.bamboo %>% filter  (plot.trt == i )
    trt_two <- cent.bamboo %>% filter  (plot.trt == j )
    trt_merge <- merge(trt_one, trt_two, by = c("Survey"))
    
    TIME1 <- dist.pt (trt_merge[1,3], trt_merge[1,4],trt_merge[1,6], trt_merge[1,7] )
    TIME2<- dist.pt (trt_merge[2,3], trt_merge[2,4],trt_merge[2,6], trt_merge[2,7] )
    TIME3 <- dist.pt (trt_merge[3,3], trt_merge[3,4],trt_merge[3,6], trt_merge[3,7] )
    TIME4 <- dist.pt (trt_merge[4,3], trt_merge[4,4],trt_merge[4,6], trt_merge[4,7] )
    TIME5 <- dist.pt (trt_merge[5,3], trt_merge[5,4],trt_merge[5,6], trt_merge[5,7] )
    TIME6 <- dist.pt (trt_merge[6,3], trt_merge[6,4],trt_merge[6,6], trt_merge[6,7] )
    TIME7 <- dist.pt (trt_merge[7,3], trt_merge[7,4],trt_merge[7,6], trt_merge[7,7] )
    TIME8 <- dist.pt (trt_merge[8,3], trt_merge[8,4],trt_merge[8,6], trt_merge[8,7] )
    TIME9 <- dist.pt (trt_merge[9,3], trt_merge[9,4],trt_merge[9,6], trt_merge[9,7] )
    
    
    dist_btwn_community <- data.frame(i, j, TIME1, TIME2, TIME3, TIME4, TIME5, TIME6, TIME7, TIME8, TIME9)
    dist_btw_trt <-rbind(dist_btw_trt, dist_btwn_community)
  }
}

col_head <- c('Treat1',"Treat2", '1',"2", "3", "4", "5", "6", "7", "8", "9")
colnames(dist_btw_trt) <- col_head

dist_btw_tidy_1 <- dist_btw_trt %>%
  gather(time, distance, 3:11) %>%
  mutate(type =rep("centroid"))
```

Man kendall trend test to detect trends in convergence or divergence between communities 

```{r man_kendall assemblage A}
###################################################################################################
# trajectory analysis 
###################################################################################################
exptime <- bamboo_community %>% select(Survey, FIRST) %>% group_by(Survey) %>% summarise_all(mean) %>%
  mutate(time = as.character(Survey))
dist_btw_tidy <- merge(dist_btw_tidy_1, exptime, by =c("time"))

mk_cent <- dist_btw_tidy %>% 
  filter(Treat1== "Coevolved") %>%
  filter(Treat2 == "Evolved residents") %>%
  mutate(time = as.numeric(time))%>% arrange(time)

mk_test1 <- mk.test(mk_cent$distance)
mk_slope <- sens.slope(mk_cent$distance)

mk_test1$estimates[3] # extract tau
mk_test1$p.value # extract pvalue from trend test 
mk_slope$estimates # extract slope 
mk_slope$p.value

# okay moving on... 
trt_level_mk_test <- dist_btw_tidy %>% 
  group_by(Treat1, Treat2) %>%  
  mutate(day = as.numeric(FIRST))%>% 
  arrange(day) %>%
  mutate(mann_kendall_tau = mk.test(distance)$estimates[3],
         mann_kendall_pvalue = mk.test(distance)$p.value,
         sens_slope = sens.slope(distance)$estimate,
         sens_pvalue = sens.slope(distance)$p.value) %>%
  group_by(Treat1, Treat2) %>%
  summarise_at(c("mann_kendall_tau", "mann_kendall_pvalue", "sens_slope", "sens_pvalue"), mean)

```
Generates convergence matrix for Assemblage A to quickly see the results and highlight significant trends 

```{r graphic_convergence assemblage A}
converg_matrix <- trt_level_mk_test %>% 
  select(Treat1, Treat2, mann_kendall_tau) %>% # arrange(Treat1) %>% arrange(Treat)%>%
  spread(Treat2, mann_kendall_tau) %>% 
  replace(., is.na(.), 0)

# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}

signif <- trt_level_mk_test %>%
mutate(signif.p = case_when((mann_kendall_pvalue < 0.05) ~ 0, 
                              (mann_kendall_pvalue > 0.05) ~ 1))%>%
  select(Treat1, Treat2, mann_kendall_pvalue, signif.p) 
conv_low <- get_lower_tri(converg_matrix)
convergence <- conv_low %>% 
  gather(Treat2, mann_kendall_tau, 2:5) %>%
  mutate(index = "Bray-Curtis")
convergence <- merge(convergence, signif, by = c("Treat1", "Treat2"))

mktrend <-convergence%>% 
  ggplot(aes(x=Treat1, y = Treat2, fill=mann_kendall_tau))+
  theme_bw(base_size =10)+
  geom_raster(aes(x=Treat1, y = Treat2, fill=mann_kendall_tau))+
  scale_fill_gradientn(limits = c(-1,1), 
                       breaks=c(-1,0,1),
                       colours=c('#d53e4f','#ffffbf','#3288bd'),
                       na.value = 'white',
                       name='Tau')+
  labs(x="", y="") +
  geom_text(aes(x=Treat1, y = Treat2, label=round(mann_kendall_tau,2)), size=3, vjust=1)+ 
  geom_text(data=subset(convergence, mann_kendall_pvalue < 0.05), 
            aes(x=Treat1, y = Treat2, label="*"), 
            size=9, color="white", vjust=0.2, hjust=-1)+ 
  geom_text(data=subset(convergence, mann_kendall_tau < -0.9),
            aes(x=Treat1, y = Treat2, label=round(mann_kendall_tau,2)),
            size=10, color="#bdbdbd",  vjust=1)+ 
  theme(axis.title.x=element_blank(),
        # axis.text.x=element_blank(),
        # axis.ticks.x=element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        legend.justification = c(1, 0),
        legend.position = c(0.5, 0.8),
        legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))+
  theme(axis.text.x = element_text( angle=90))+
  guides(fill=FALSE)
#jpeg(filename="convergence_AssemblageA.jpeg", width=90, height=120, units="mm", bg="white", res=300)
mktrend
#dev.off()

```
Lets look at all the comparisons

```{r all comparisons}

#jpeg(filename="convergence_allcomparions_AssemblageA.jpeg", width=270, height=270, units="mm", bg="white", res=300)
dist_btw_tidy %>%
  #filter(!(Treat1==Treat2))%>%
  ggplot() +
  geom_point(aes(x= as.numeric(FIRST), y = distance), size=4)+
   geom_smooth(method='lm',aes(x= as.numeric(FIRST), y = distance), se=FALSE, col='black')+
  labs(x="days", y="distance between communities")+
  facet_wrap(~Treat1*Treat2, scales = 'free_y')+
  rita_theme+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        strip.text = element_text(size = 15))
#dev.off()
```

~~~~~~~~~~~~~~~~~~ Now doing trajectory analysis for residents only, and including the uninvaded control into trajectory analysis

*** note that the invader was removed from the following analysis 

```{r NMDS assemblage A resident}
# create community matrix and envi data matrix 
bamb_mat<-bamboo_community_res[,c(6:9,11)] # matrix of protist species
bamb_mat_treat<-bamboo_community_res[,c(1:5, 14:18)] # matrix of treatments

 # global nmds
set.seed(123456789)
 ordO_pt<-metaMDS(bamb_mat, distance="bray", trymax = 500, autotransform=FALSE)
 ordO_pt

 #extract info for plotting
 #add species names to ordination
  species <- c("Blepharisma", "Spirostomum", "Euplotes patella", "Lecane", #"Prorodon",
               "Paramecium bursaria")
  hcoordO<-as.data.frame(scores(ordO_pt, display="sites"))#extracts coordinates for plot
  pcoordO<-scores(ordO_pt, display="species")#extracts coordinates for parasite vectors
  pcoordO <-as.data.frame(pcoordO)
  pcoordO$species <- species
 
  #make dataframe with NMDS coordinates and treatment info 
 bamboo_all_resident <- bind_cols(bamb_mat_treat, hcoordO)
 #write.csv(bamboo_all, "NMDS_coordinates_bamboo_assemblage.csv")
 
 cent.bamboo.res <- bamboo_all_resident %>%
  group_by(plot.trt, Survey) %>%
  summarise_at(c("NMDS1", 'NMDS2'), mean)
```
This is the centroid graphic for the residents only 

```{r graphic_nmds_centroid assemblage A resident only}
topp<-max(cent.bamboo.res[,3:4]) #determines maximum and minimum values for the plot axes
bott<-min(cent.bamboo.res[,3:4]) #I draw from both data sets because I make the graph sqaure and

# #jpeg(filename="AssemblageA_trajectory_residentonly.jpeg", width=180, height=180, units="mm", bg="white", res=300)
ggplot(data= cent.bamboo.res, aes(x= NMDS1, y=NMDS2, group=plot.trt))+
  geom_hline(yintercept = 0, lty=2, color="grey") +
  geom_vline(xintercept = 0, lty=2, color="grey") +
  geom_path(color="#525252", lwd=1)+
  geom_point( size =6, aes(x= NMDS1, y=NMDS2, fill= plot.trt), 
              pch=21)+
  xlab("NMDS 1") +
  ylab("NMDS 2") +
  xlim(c( bott, topp)) +
  ylim(c( bott, topp)) +
  scale_fill_manual(values =colorpal2)+
  geom_text( aes(x=NMDS1,y=NMDS2, label=Survey), fontface =2, colour = "black" , size=5) + # plant species label
  geom_text(aes(x=NMDS1,y=NMDS2, label=Survey), fontface =1, colour = "black" , size=4.5) + # plant species label
 # geom_text(data=pcoordO, aes(x=NMDS1,y=NMDS2, label=species), fontface =2, colour = "black" , size=2) + # species label
  theme_pubr() +
  theme(legend.text = element_text(size=8), legend.box = "horizontal",
        legend.title = element_text(size=9, face="bold"),
        legend.position=c(0.13,0.85)) +
  guides(colour = guide_legend(override.aes = list(size=70)))+
  theme(plot.title = element_text(hjust = 0.5, size=15, face="bold")) + # guides(shape=FALSE)+
facet_wrap(~plot.trt,
             ncol = 2, nrow = 3)+
  guides(fill=FALSE, color=FALSE)+
  ggtitle("Assemblage A")+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        strip.text = element_text(size = 15)) # add black square panel around graphic
# #dev.off()
```

```{r trajectory resident only}
###################################################################################################
# distance between treatment centroids: how dissimilar are the trajectories btwn groups 
###################################################################################################
dist.pt <- function(x1, y1, x2, y2) sqrt((x2-x1)^2 + (y2-y1)^2) # finds distance between points 

cent.bamboo.res <- bamboo_all_resident %>%
  group_by(plot.trt, Survey) %>%
  summarise_at(c("NMDS1", 'NMDS2'), mean) 

trt_list <- unique(cent.bamboo.res$plot.trt)
dist_btw_trt_resident <- data.frame()

for (i in trt_list){
  for(j in trt_list){
    
    treat <- i
    treat2 <- j 
    
    trt_one <- cent.bamboo.res %>% filter  (plot.trt == i )
    trt_two <- cent.bamboo.res %>% filter  (plot.trt == j )
    trt_merge <- merge(trt_one, trt_two, by = c("Survey"))
    
    TIME1 <- dist.pt (trt_merge[1,3], trt_merge[1,4],trt_merge[1,6], trt_merge[1,7] )
    TIME2<- dist.pt (trt_merge[2,3], trt_merge[2,4],trt_merge[2,6], trt_merge[2,7] )
    TIME3 <- dist.pt (trt_merge[3,3], trt_merge[3,4],trt_merge[3,6], trt_merge[3,7] )
    TIME4 <- dist.pt (trt_merge[4,3], trt_merge[4,4],trt_merge[4,6], trt_merge[4,7] )
    TIME5 <- dist.pt (trt_merge[5,3], trt_merge[5,4],trt_merge[5,6], trt_merge[5,7] )
    TIME6 <- dist.pt (trt_merge[6,3], trt_merge[6,4],trt_merge[6,6], trt_merge[6,7] )
    TIME7 <- dist.pt (trt_merge[7,3], trt_merge[7,4],trt_merge[7,6], trt_merge[7,7] )
    TIME8 <- dist.pt (trt_merge[8,3], trt_merge[8,4],trt_merge[8,6], trt_merge[8,7] )
    TIME9 <- dist.pt (trt_merge[9,3], trt_merge[9,4],trt_merge[9,6], trt_merge[9,7] )
    
    
    dist_btwn_community <- data.frame(i, j, TIME1, TIME2, TIME3, TIME4, TIME5, TIME6, TIME7, TIME8, TIME9)
    dist_btw_trt_resident <-rbind(dist_btw_trt_resident, dist_btwn_community)
  }
}

col_head <- c('Treat1',"Treat2", '1',"2", "3", "4", "5", "6", "7", "8", "9")
colnames(dist_btw_trt_resident) <- col_head

dist_btw_tidy_1_resident <- dist_btw_trt_resident %>%
  gather(time, distance, 3:11) %>%
  mutate(type =rep("centroid"))
```

```{r man_kendall assemblage A resident only}
###################################################################################################
# trajectory analysis 
###################################################################################################
exptime <- bamboo_community_res %>% select(Survey, FIRST) %>% group_by(Survey) %>% summarise_all(mean) %>%
  mutate(time = as.character(Survey))
dist_btw_tidy_1_resident <- merge(dist_btw_tidy_1_resident, exptime, by =c("time"))

mk_cent <- dist_btw_tidy_1_resident %>% 
  filter(Treat1== "Coevolved") %>%
  filter(Treat2 == "Evolved residents") %>%
  mutate(time = as.numeric(time))%>% arrange(time)

mk_test1 <- mk.test(mk_cent$distance)
mk_slope <- sens.slope(mk_cent$distance)

mk_test1$estimates[3] # extract tau
mk_test1$p.value # extract pvalue from trend test 
mk_slope$estimates # extract slope 
mk_slope$p.value

# okay moving on... 
trt_level_mk_test_resident <- dist_btw_tidy_1_resident %>% 
  group_by(Treat1, Treat2) %>%  
  mutate(day = as.numeric(FIRST))%>% 
  arrange(day) %>%
  mutate(mann_kendall_tau = mk.test(distance)$estimates[3],
         mann_kendall_pvalue = mk.test(distance)$p.value,
         sens_slope = sens.slope(distance)$estimate,
         sens_pvalue = sens.slope(distance)$p.value) %>%
  group_by(Treat1, Treat2) %>%
  summarise_at(c("mann_kendall_tau", "mann_kendall_pvalue", "sens_slope", "sens_pvalue"), mean)

```


```{r graphic_convergence assemblage A resident only}
converg_matrix <- trt_level_mk_test_resident %>% 
  select(Treat1, Treat2, mann_kendall_tau) %>% # arrange(Treat1) %>% arrange(Treat)%>%
  spread(Treat2, mann_kendall_tau) %>% 
  replace(., is.na(.), 0)

# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
}

signif <- trt_level_mk_test_resident %>%
mutate(signif.p = case_when((mann_kendall_pvalue < 0.05) ~ 0, 
                              (mann_kendall_pvalue > 0.05) ~ 1))%>%
  select(Treat1, Treat2, mann_kendall_pvalue, signif.p) 
conv_low <- get_lower_tri(converg_matrix)
convergence <- conv_low %>% 
  gather(Treat2, mann_kendall_tau, 2:5) %>%
  mutate(index = "Bray-Curtis")
convergence_resdient <- merge(convergence, signif, by = c("Treat1", "Treat2"))

mktrend <-convergence_resdient%>% 
  ggplot(aes(x=Treat1, y = Treat2, fill=mann_kendall_tau))+
  theme_bw(base_size =10)+
  geom_raster(aes(x=Treat1, y = Treat2, fill=mann_kendall_tau))+
  scale_fill_gradientn(limits = c(-1,1), 
                       breaks=c(-1,0,1),
                       colours=c('#d53e4f','#ffffbf','#3288bd'),
                       na.value = 'white',
                       name='Tau')+
  labs(x="", y="") +
  geom_text(aes(x=Treat1, y = Treat2, label=round(mann_kendall_tau,2)), size=3, vjust=1)+ 
  geom_text(data=subset(convergence_resdient, mann_kendall_pvalue < 0.05), 
            aes(x=Treat1, y = Treat2, label="*"), 
            size=9, color="white", vjust=0.2, hjust=-1)+ 
  geom_text(data=subset(convergence_resdient, mann_kendall_tau < -0.9),
            aes(x=Treat1, y = Treat2, label=round(mann_kendall_tau,2)),
            size=10, color="#bdbdbd",  vjust=1)+ 
  theme(axis.title.x=element_blank(),
        # axis.text.x=element_blank(),
        # axis.ticks.x=element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        legend.justification = c(1, 0),
        legend.position = c(0.5, 0.8),
        legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                               title.position = "top", title.hjust = 0.5))+
  theme(axis.text.x = element_text( angle=90))+
  guides(fill=FALSE)
#jpeg(filename="convergence_AssemblageA_residentonly.jpeg", width=90, height=120, units="mm", bg="white", res=300)
mktrend
#dev.off()

```

~~~~~~~~~~~~~~~~Population level analysis using mvabund mixed models 


```{r mvabund model}
library(mvabund)
species_list <- list(bamboo_community[c(6:10,12)])
treatment <- list(bamboo_community[c(3,19)])
block <- list(bamboo_community[4])

mv.list <-list(species_1 = species_list,
              treatment_1 = treatment, 
              block_1 = block)
attach(mv.list)
species.data <- mvabund(species_1)
is.mvabund(species.data)

#constrain the permutation within a jar by designating shuffletset; 
#nrow = within the abundance data 
#control within jar= block  
permID <- shuffleSet(n=nrow(species.data), nset=999, control=how(block=bamboo_community$JAR))

#set reference level
bamboo_community <-bamboo_community %>% mutate(invader_evol=as.factor(invader_evol),
                            resident_evol = as.factor(resident_evol),
                            plot.trt =as.factor(plot.trt))
bamboo_community$invader_evol <- relevel(bamboo_community$invader_evol, ref='naive')
bamboo_community$resident_evol <- relevel(bamboo_community$resident_evol , ref='naive')
bamboo_community$plot.trt <- relevel(bamboo_community$plot.trt , ref='Naive invasion')

#community composition analysis
  mod1 <- manyglm(species.data ~ (FIRST)* invader_evol *resident_evol, 
                  data= bamboo_community,
                  family='negative.binomial',
                  show.residuals=T,
                  show.coef = T)
  
    mod1.1 <- manyglm(species.data ~ (FIRST)* plot.trt, 
                  data= bamboo_community,
                  family='negative.binomial',
                  show.residuals=T,
                  show.coef = T)
    
plot.manyglm(mod1)
coef(mod1)
summary(mod1)
#jpeg(filename="assemblageA_glm_coeff.jpeg", width=220, height=240, units="mm", bg="white", res=300)
#coefplot(mod1)
#dev.off()
#summary(mod1)
new <- data.frame(bamboo_community[c(15,18,20)])
mod.pred <- predict.manyglm(mod1.1, new, type="response") 
mod.pred.predictor <- predict.manyglm(mod1.1, new) 
plot.mod <- cbind(new, mod.pred)

#get SE for fitted line
mod.pred.se <-predict.manyglm(mod1, new, type='response', se.fit=TRUE)
se <- mod.pred.se$se.fit %>% as.data.frame()
sp <- c("Blepharisma", "Spirostomum","Euplotes_patella","Euplotes_daidaleos1","Lecane","Paramecium_bursaria1") 
colnames(se) <- sp

#permutation for p values based on repeated measures 
set.seed(12345)
mod.out <- anova(mod1, bootID=permID, p.uni = 'adjusted', test = 'LR' )
printCoefmat(mod.out$uni.p)
mod.out

#get table of p values for univariate tests for each species 
AssembA.table <- as.data.frame(mod.out$uni.test)
```

Plot model predictions for multivariate glms 
Note we need to back transform model fit 

```{r model prediction plot}
#back transformed model fitted line 
plot.mod.2 <-plot.mod %>% 
  gather(species, abundance, c(4:9))
plot.se <- se %>%
  gather(species, se, 1:6) %>% 
  select(-c(species))
#add SE measure to model fitted line 
plot.mod.2$se <- plot.se$se
plot.mod.2$plot.trt <- bamboo_community$plot.trt
  
#get means for plotting 
mean.abund <-bamboo_community %>% 
  ungroup()%>%
  gather(species, abundance, c(6:10,12)) %>%
  group_by(species, plot.trt,Survey)%>%
  mutate(mean_abun=mean(abundance),
         sd_abun= sd(abundance),
         n_abun = 5) %>%
  ungroup()%>%
  group_by(species, plot.trt,Survey)%>%
  mutate(se_abun = sd_abun / sqrt(n_abun),
         lower_ci = mean_abun - qt(1 - (0.05 / 2), n_abun - 1) * se_abun,
         upper_ci = mean_abun + qt(1 - (0.05 / 2), n_abun - 1) * se_abun)

dodge <- position_dodge(width = 2)

species_names <- list(
'Blepharisma' = 'B. americanum',
'Euplotes_daidaleos1' = 'E. daidaleos',
'Euplotes_patella' = 'E. patella',
'Lecane' = 'Lecane sp.',
'Paramecium_bursaria1' = 'P. bursaria',
'Spirostomum' = 'S. teres'
)

treatment_labeller <- function(variable,value){
  return(species_names[value])
}

spp.abund.graph <-ggplot(aes(x=FIRST, y = (abundance), group=plot.trt, col=plot.trt), data= plot.mod.2) +
  geom_line(aes(x=FIRST, y = (abundance), group=plot.trt, col=plot.trt),
            lwd=2, data=plot.mod.2)+
  geom_ribbon(aes(ymin = (abundance-se), ymax = (abundance+se),fill=plot.trt),
              data = plot.mod.2,alpha = 0.5, colour = NA)+
  geom_point(aes(x=FIRST, y = mean_abun, fill=plot.trt), 
             data=mean.abund, pch=21, col='black', size=3, position=dodge)+
  rita_theme+
  facet_wrap(~species, scales='free_y', labeller = treatment_labeller, ncol=2) + 
  scale_color_manual(values=colorpal)+
  scale_fill_manual(values = colorpal)+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        strip.text = element_text(size = 15))+
  theme(legend.title = element_text(size=8.5, face="bold"),
        legend.text = element_text(size=14))+ 
  theme(legend.position="top")+
  guides(color=guide_legend(""), fill=FALSE)+
  labs(x="time (days)", y ='abundance')

#jpeg(filename="assemblageA_species_level.jpeg", width=220, height=240, units="mm", bg="white", res=300)
spp.abund.graph
#dev.off()


```
~~~~~~~~~~~ resident species only comparison 

```{r mvabund model residents only}
library(mvabund)
species_list <- list(bamboo_community_res[c(6:9,11)])
treatment <- list(bamboo_community_res[c(3,19)])
block <- list(bamboo_community_res[4])

mv.list.res <-list(species_1 = species_list,
              treatment_1 = treatment, 
              block_1 = block)
attach(mv.list.res)
species.data.res <- mvabund(species_1)
is.mvabund(species.data.res)

#constrain the permutation within a jar by designating shuffletset; 
#nrow = within the abundance data 
#control within jar= block  
permID <- shuffleSet(n=nrow(species.data.res),
                     nset=999, 
                     control=how(block=bamboo_community_res$JAR))

#set reference level
bamboo_community_res <-bamboo_community_res %>% 
  mutate(invader_evol=as.factor(invader_evol),
         resident_evol = as.factor(resident_evol),
         invasion = as.factor(invasion))
bamboo_community_res$invader_evol <- relevel(bamboo_community_res$invader_evol, ref='naive')
bamboo_community_res$resident_evol <- relevel(bamboo_community_res$resident_evol , ref='naive')
bamboo_community_res$invasion <- relevel(bamboo_community_res$invasion , ref='uninvaded')

#community composition analysis
mod1.res <- manyglm(species.data.res ~ FIRST *invasion,
                    data=bamboo_community_res,
                    family='negative.binomial',
                    show.residuals=T)
mod1.res.plot <- manyglm(species.data.res ~ FIRST *plot.trt,
                    data=bamboo_community_res,
                    family='negative.binomial',
                    show.residuals=T)
plot.manyglm(mod1.res)
summary(mod1.res)
coef(mod1.res)
#extract info to replot model results
new <- data.frame(bamboo_community_res[c(14,16)])
mod.pred.res <-predict(mod1.res.plot, new)
plot.mod.res <- cbind(new, mod.pred.res)

#get SE for fitted line
mod.pred.se.res <-predict(mod1.res.plot, new, se=TRUE)
se <- mod.pred.se.res$se.fit %>% as.data.frame()
sp <- c("Blepharisma", "Spirostomum","Euplotes_patella","Lecane","Paramecium_bursaria1") 
colnames(se) <- sp

#permutation for p values based on repeated measures 
set.seed(12345)
mod.out.res <- anova.manyglm(mod1.res, bootID=permID,  p.uni = 'adjusted', test = 'LR')
printCoefmat(mod.out.res$uni.p)
mod.out.res

#get table of p values for univariate tests for each species 
AssembA.table.resident <- as.data.frame(mod.out.res$uni.p)
```

Graphic of species change

```{r resident species populations}

#back transformed model
plot.mod.2.res <-plot.mod.res %>% 
  gather(species, abundance, c(3:7))
plot.se <- se %>%
  gather(species, se, 1:5) %>% 
  select(-c(species))
plot.mod.2.res$se <- plot.se$se

species_names <- list(
'Blepharisma' = 'B. americanum',
'Euplotes_daidaleos1' = 'E. daidaleos',
'Euplotes_patella' = 'E. patella',
'Lecane' = 'Lecane sp.',
'Paramecium_bursaria1' = 'P. bursaria',
'Spirostomum' = 'S. teres'
)

treatment_labeller <- function(variable,value){
  return(species_names[value])
}

#get means for plotting 
mean.abund.res <-bamboo_community_res %>% 
  ungroup()%>%
  gather(species, abundance, c(6:9,11)) %>%
  group_by(species, plot.trt,Survey)%>%
  mutate(mean_abun=mean(abundance),
         sd_abun= sd(abundance),
         n_abun = 5) %>%
  ungroup()%>%
  group_by(species, plot.trt,Survey)%>%
  mutate(se_abun = sd_abun / sqrt(n_abun),
         lower_ci = mean_abun - qt(1 - (0.05 / 2), n_abun - 1) * se_abun,
         upper_ci = mean_abun + qt(1 - (0.05 / 2), n_abun - 1) * se_abun)

dodge <- position_dodge(width = 2)
  
sp.abund.graphic.resident <-ggplot(aes(x=FIRST, y = exp(abundance), group=plot.trt, col=plot.trt), data= plot.mod.2.res) +
  geom_ribbon(aes(ymin = exp(abundance-se), ymax = exp(abundance+se),fill=plot.trt),
              data = plot.mod.2.res,alpha = 0.5, colour = NA)+
    geom_line(aes(x=FIRST, y = exp(abundance), group=plot.trt, col=plot.trt),
            lwd=2, data=plot.mod.2.res)+
  geom_point(aes(x=FIRST, y = mean_abun, fill=plot.trt), 
             data=mean.abund.res, pch=21, col='black', size=3, position=dodge)+
  rita_theme+
  facet_wrap(~species, scales='free_y', labeller = treatment_labeller, ncol=2) + 
  scale_color_manual(values=c('#d7191c','#fdae61','#abdda4','#2b83ba', '#737373'))+
  scale_fill_manual(values = colorpal2)+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        strip.text = element_text(size = 15))+
  theme(legend.title = element_text(size=8.5, face="bold"),
        legend.text = element_text(size=14))+ 
  theme(legend.position="top")+
  guides(color=guide_legend(""), fill=FALSE)+
  labs(x="time (days)", y ='abundance')

#jpeg(filename="assemblageA_species_level_residents.jpeg", width=180, height=180, units="mm", bg="white", res=300)
sp.abund.graphic.resident
#dev.off()
```
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ DRIFT ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
To assess drift within a treatment, I calculated the distance from the centroid for each community within a treatment group.   

```{r drift}
###################################################################################################
# calculate distance between communities and their centroid
dist.pt <- function(x1, y1, x2, y2) sqrt((x2-x1)^2 + (y2-y1)^2) # finds distance between points 

jar_list <- unique(bamboo_community$JAR)

bamboo_nmds <- merge(bamboo_all, cent.bamboo, by =c ("plot.trt", "Survey"))

dis_centroid <- data.frame()

for (i in jar_list){

    jar <- i
    
   # trt_merge <- bamboo_nmds %>% filter(JAR=="B1R1") %>% arrange(Survey)
    trt_merge <- bamboo_nmds %>% 
      filter  (JAR== i)%>%
      arrange(Survey)
    
    TIME1 <- dist.pt (trt_merge[1,11], trt_merge[1,12],trt_merge[1,13], trt_merge[1,14] )
    TIME2 <- dist.pt (trt_merge[2,11], trt_merge[2,12],trt_merge[2,13], trt_merge[2,14] )
    TIME3 <- dist.pt (trt_merge[3,11], trt_merge[3,12],trt_merge[3,13], trt_merge[3,14] )
    TIME4 <- dist.pt (trt_merge[4,11], trt_merge[4,12],trt_merge[4,13], trt_merge[4,14] )
    TIME5 <- dist.pt (trt_merge[5,11], trt_merge[5,12],trt_merge[5,13], trt_merge[5,14] )
    TIME6 <- dist.pt (trt_merge[6,11], trt_merge[6,12],trt_merge[6,13], trt_merge[6,14] )
    TIME7 <- dist.pt (trt_merge[7,11], trt_merge[7,12],trt_merge[7,13], trt_merge[7,14] )
    TIME8 <- dist.pt (trt_merge[8,11], trt_merge[8,12],trt_merge[8,13], trt_merge[8,14] )
    TIME9 <- dist.pt (trt_merge[9,11], trt_merge[9,12],trt_merge[9,13], trt_merge[9,14] )
    
    dist_btwn_jar <- data.frame(i,trt_merge$plot.trt, 
                                  TIME1, TIME2, TIME3, TIME4, 
                                TIME5, TIME6, TIME7, 
                                  TIME8, TIME9)
    dis_centroid <-rbind(dis_centroid, dist_btwn_jar)

}

col_head <- c('JAR',
              "plot.trt", 
              '0', '14', '30', '36', '51', '56', '63', '70', '76')
colnames(dis_centroid) <- col_head

distance_centroid <- dis_centroid %>%
  group_by(JAR, plot.trt) %>%
  summarise_all(mean) %>%
  gather(time, Distance,3:11)%>%
  drop_na()
```

Visualize distance from centroid over time in the graphic below

```{r drift graphic}

distance_centroid %>% ggplot()+
  geom_boxplot(aes(x=as.factor(time), y = Distance, fill=plot.trt))+ 
 # facet_wrap(~plot.trt)+
  labs(x="time", y = "distance to centroid")+
  rita_theme+
  guides(fill=FALSE)+
  scale_fill_manual(values=colorpal)+
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1),
        strip.text = element_text(size = 15)) # add black square panel around graphic

#jpeg(filename="driftmodel_AssemblageA.jpeg", width=130, height=100, units="mm", bg="white", res=300)
distance_centroid %>% ggplot(aes(x=(time), y = Distance, group=plot.trt))+
  geom_point(aes(x=(time), y = Distance, fill=plot.trt), pch=21,
             position=position_jitterdodge(dodge.width=0.5,  jitter.width = 0.25), size=2, alpha=0.5)+ 
  geom_smooth(aes(x=(time), y = Distance, color=plot.trt, fill=plot.trt), method='lm', se=FALSE, lwd=2)+ 
  scale_color_manual(values=colorpal)+
  scale_fill_manual(values=colorpal)+
  # facet_wrap(~plot.trt)+
  labs(x="time", y = "distance to centroid")+
  rita_theme+
  theme(legend.title = element_text(size=8.5, face="bold"),
        legend.text = element_text(size=8))+ 
  guides(color=guide_legend(""), fill=FALSE)+
  theme(legend.position="top")
#dev.off()
```
Fit a mixed model

```{r drift model}
library(lmerTest)

cent.model <-lmer((Distance) ~ as.numeric(time) * plot.trt+ (1|JAR), data= distance_centroid) #random intercept
plot(cent.model)
summary(cent.model)
```
